Wall-to-wall mapping of peat depth from Lidar terrain and airborne radiometrics in Norwegian landscapes

### 2024-11-03 ms/ms.Rmd
ms: results - pdps

>>>>

Many of the the most important predictors in the best performing models showed non-monotonic effects on peat depth (Fig. \@ref(fig:pdps)).
At Ørskogfjellet for example, increasing *elevation* was predictive of deeper peat up to about 75 meters above sea level, after which a further increase was predictive of shallower peat.
At Skrimfjella the partial dependence on *elevation* had the opposite shape, with the shallowest peats predicted at intermediate elevations, around 350 meters above sea level.
*TWI50m* at Ørskogfjellet and *DTW4000* and *radU* at Skrimfjella were other predictors that showed considerable fluctuations in their predictive effects on peat depth.
The radiometric predictor in particular displayed an idiosyncratic effect, with a marked dip in predicted depth at intermediate values of *radU*.
On the other hand, the partial effects of some important predictors were more straightforward.
The partial dependence on *MRVBF* was quite similar across sites, with the deepest peats predicted in the very flattest valley bottoms.
Also, *TWI5m* and *DTW2500* at Skrimfjella showed monotonically positive and negative predictive effects, respectively.

Individual conditional expectation lines indicated some interactions between predictors (Fig. \@ref(fig:pdps)).
For example, the magnitude of the increase in depth with *elevation* that the model expected at Ørskogfjellet was different for different observations; some depth predictions increased by only \unit{40\,cm} while others increased by more than \unit{100\,cm}, over the same elevation gain.
Similarly, ICE lines of *radU* at Skrimfjella were non-parallel, with some locations showing monotonically increasing peat depth predictions with uranium concentration (unlike the average effect).
Nevertheless, most ICE lines were generally parallel -- indicating that the average effects of the predictors were good representations of their overall effects.

```{r pdps, echo = FALSE, fig.cap = "Partial dependence plots of the six most important variables in the best-performing models at Ørskogfjellet (a) and Skrimfjella (b). The average effect of the predictor on the outcome (red line) overlays the variation in the effect across observations (grey lines)."}
knitr::include_graphics("figures/partial_dependence.pdf")
```

### 2024-11-03 ms/ms.tex
ms: results - pdps

>>>>

Many of the the most important predictors in the best performing models showed non-monotonic effects on peat depth (Fig. \ref{fig:pdps}).
At Ørskogfjellet for example, increasing \emph{elevation} was predictive of deeper peat up to about 75 meters above sea level, after which a further increase was predictive of shallower peat.
At Skrimfjella the partial dependence on \emph{elevation} had the opposite shape, with the shallowest peats predicted at intermediate elevations, around 350 meters above sea level.
\emph{TWI50m} at Ørskogfjellet and \emph{DTW4000} and \emph{radU} at Skrimfjella were other predictors that showed considerable fluctuations in their predictive effects on peat depth.
The radiometric predictor in particular displayed an idiosyncratic effect, with a marked dip in predicted depth at intermediate values of \emph{radU}.
On the other hand, the partial effects of some important predictors were more straightforward.
The partial dependence on \emph{MRVBF} was quite similar across sites, with the deepest peats predicted in the very flattest valley bottoms.
Also, \emph{TWI5m} and \emph{DTW2500} at Skrimfjella showed monotonically positive and negative predictive effects, respectively.

Individual conditional expectation lines indicated some interactions between predictors (Fig. \ref{fig:pdps}).
For example, the magnitude of the increase in depth with \emph{elevation} that the model expected at Ørskogfjellet was different for different observations; some depth predictions increased by only \unit{40\,cm} while others increased by more than \unit{100\,cm}, over the same elevation gain.
Similarly, ICE lines of \emph{radU} at Skrimfjella were non-parallel, with some locations showing monotonically increasing peat depth predictions with uranium concentration (unlike the average effect).
Nevertheless, most ICE lines were generally parallel -- indicating that the average effects of the predictors were good representations of their overall effects.

\begin{figure}
\centering
\includegraphics{figures/partial_dependence.pdf}
\caption{\label{fig:pdps}Partial dependence plots of the six most important variables in the best-performing models at Ørskogfjellet (a) and Skrimfjella (b). The average effect of the predictor on the outcome (red line) overlays the variation in the effect across observations (grey lines).}
\end{figure}
